---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: Dissolved Organic Carbon in Lakes from the North Temperate Region
subtitle: https://github.com/cwatson1013/Env_Data_Analysis_Final_Proj.git
author: Caroline Watson
abstract: "Dissolved organic carbon (DOC) is a significant part of the carbon cycle. DOC can come from within the lake, such as from decaying organisms or plant matter, or from the catchment area of the lake. The catchment area soils can influence the amount of DOC that enters each lake. The purpose of this study is to see if there is a relationship between depth and DOC and if so, if there is a seasonal relationship. This study looked at lakes that were part of the North Temperate Region located in Wisconsin, USA. DOC was found to not be normally distributed, so a two-way ANOVA with an interaction test was conducted. The results found that depth and lake name were significant. A non-parametric test was run on Peter, Paul, and West Long lakes to determine trends and changepoints in the data. Peter and West Long lake had trends and changepoints, but Paul lake did not. Overall, it was found that the depth and DOC do not have a relationship and a seasonal relationship between change in depth and DOC could not be determined with the data."
fontsize: 12pt
mainfont: Times New Roman
---

<Information in these brackets are used for annotating the RMarkdown file. They will not appear in the final version of the PDF document>

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

<Setup the global options for the R chunks in your document>

```{r setup, warning=FALSE, message=FALSE, echo=TRUE, eval=TRUE}

knitr::opts_chunk$set(echo = TRUE, eval = TRUE,
                      cache = TRUE, fig.pos = "H", warning = FALSE, message = FALSE)

# Set your working directory
setwd("/Users/carolinewatson/Documents/Spring 2019/Environmental Data Analytics/Final_Project/Data/Code")

# Load your packages
suppressMessages(library(tidyverse))
library(ggplot2)
library(leaflet)
library(dplyr)
library(RColorBrewer)
library(viridis)
library(knitr)
library(kableExtra)
library(lubridate)
library(lme4)
library(nlme)
library(trend)
library(gridExtra)
library(multcomp)

# Set your ggplot theme
caroline_theme <- theme_classic(base_size = 16) + 
  theme(axis.text = element_text(color = "black"), 
        legend.position = "right")

theme_set(caroline_theme)
```


# Research Question and Rationale

<Paragraph detailing the rationale for your analysis. What is the significant application and/or interest in this topic? Connect to environmental topic(s)/challenge(s).>

The rationale for this analysis is that there typically is a relationship between dissolved organic carbon and depth. There is also typically a relationship between land area surrounding lakes that have high amounts of organic soils usually deposit large amounts of dissolved organic carbon into lakes. Dissolved organic carbon is an important part of the carbon cycle and supplies nutrients for some organisms. Most DOC is natural, but high amounts can indicate human influence, such as land surrounding the lake that is high in organic amount. 

<Paragraph detailing your research question(s) and goals. What do you want to find out? Include a sentence (or a few) on the dataset you are using to answer this question - just enough to give your reader an idea of where you are going with the analysis.>

This study aims to find out whether there is a relationship between dissolved organic carbon (DOC) and depth. If there is a relationship between these two variables, I want to see if this relationship varies seasonally. This dataset contains various parameter measurements for different lakes in the North Temperate Region in Wisconsin, USA that were part of the Long Term Ecological Research station. The measured parameters include temperature, depth, dissolved organic carbon, dissolved inorganic carbon, particulate organic matter and others.

\newpage

# Dataset Information

<Information on how the dataset for this analysis were collected, the data contained in the dataset, and any important pieces of information that are relevant to your analyses. This section should contain much of same information as the README file for the dataset but formatted in a way that is more narrative.>

<Add a table that summarizes your data structure. This table can be made in markdown text or inserted as a `kable` function in an R chunk. If the latter, do not include the code used to generate your table.>

The data for this analysis were collected from June 3, 1984 - August 17, 2016. Dissolved inorganic carbon (DIC), dissolved organic carbon (DOC), particulate organic matter, partial pressure of CO2, and absorbance at 440nm were all parameters that were measured. Peter, Paul, West Long, East Long, Tuesday, and Crampton lakes were all sampled. The samples for dissolved organic carbon and absorbance were collected from the epilimnion, metalimnion, hypolimnion and PML (Pooled Mixed Layer). It is important to note that there were no numeric depths corresponding to DOC or absorbance. The samples for inorganic carbon were taken at 100%, 50%, 25%, 10%, 5%, and 1% of the surface irridance, with some samples being collected from the hypolimnion as well. Two types of samples for partial pressure of CO2 were obtained: one was from the air, the other from the lake. The air sample was taken at 2 meters above the lake, and the water sample was taken right below the surface of the lake.  from 2 meters above the surface of the lake,  Table 1 shows a summary of all the measurements in the carbon dataset. The frequency of the sampling varied. 
 
```{r}

#reading in data file
carbon.data <- read.csv("../Raw/NTL-LTER_Lake_Carbon_Raw.csv")

#structure of data frame
carbon.data.summary <- summary(carbon.data)
```


```{r, echo=FALSE, eval=TRUE, tbls="Summary of NTL-LTER Lake Carbon Data", results="asis"}
#summary of data structure
kable(carbon.data.summary, caption = "Summary of Carbon Data from NTL-LTER Lakes in Wisconsin") %>% kable_styling(latex_options = c("hold_position", "striped", "scale_down")) %>% kableExtra::landscape() 

```


\newpage

# Exploratory Data Analysis and Wrangling

<Include R chunks for 5+ lines of summary code (display code and output), 3+ exploratory graphs (display graphs only), and any wrangling you do to your dataset(s).> 

```{r}
#class of sampledate column
class(carbon.data$sampledate)

#converting sampledate to a date in R
carbon.data$sampledate <- as.Date(carbon.data$sampledate, format = "%m/%d/%y")

#checking class of sampledate
class(carbon.data$sampledate)

#summary of the dataset
head(carbon.data)
summary(carbon.data)
colnames(carbon.data)
dim(carbon.data)

#renaming columns
colnames(carbon.data)[1:5] <- c("Lake.ID", "Lake.Name", "Year", "Day.Number", "Date")
```

The dataset was imported and the date column was formatted as a date in R. A summary of the dataset was run to understand what the dataset contained. The head of the dataset was viewed as well as the dimensions and column names. Some column names were changed to make it easier to identify them.

```{r}
#wrangling dataset
carbon.data.skinny <- carbon.data %>%
  filter(depth_id == -2 | depth_id == -1 | depth_id == 7) %>%
  filter(depth == "Hypolimnion" | depth == "Epilimnion" | depth == "PML" |
           depth == "Metalimnion") %>%
  mutate(depth = factor(depth, levels = c("Epilimnion", "Metalimnion",
                                          "PML", "Hypolimnion"))) %>%
  filter(is.na(doc) == FALSE) %>%
  dplyr::select(Lake.ID:depth_id, doc)

#saving processed file to the processed folder
write.csv(carbon.data.skinny, 
          file = "../Processed/NTL-LTER_Lake_Carbon_Processed.csv")
```

The dataset was wrangled by filtering for depth id and depth description. DOC measurements were only taken at the epilimnion, metalimnion, PML, and hypolimnion, so the dataset needed to be filtered by these options. This study focused mainly on DOC and depth, so columns such as total particulate carbon, or irridance, were not needed. Columns that were needed were selected when the data was wrangled. The depths were also mututated as factor levels so that future analyses would present the depths from shallowest to deepest (epilimnion to hypolimnion).

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig1}QQplot showing whether data for DOC is normally distributed."}
#qqplot of doc to see if there is a normal relationship
qqnorm(carbon.data.skinny$doc) 
qqline(carbon.data.skinny$doc)
```

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig2}Log QQplot of dissolved organic carbon data."}
#qqplot of log(doc) to see if this is better than the regular qqplot
qqnorm(log(carbon.data.skinny$doc))
qqline(log(carbon.data.skinny$doc))
```

\autoref{fig:fig1} is a qqplot that shows whether the dissolved organic carbon data has a normal distribution. From \autoref{fig:fig1}, it is clear that the dissolved organic carbon data do not follow a normal distribution. Becuase the data are not normally distributed, a log qqplot was run to see if log transforming the data would help the data look more normally distributed. \autoref{fig:fig2} shows the dissolved organic carbon data log transformed. From \autoref{fig:fig2} it is clear that dissolved organic carbon is not normally distributed, but log transforming the data makes the data more normally distributed than not log transforming the data. 

```{r, eval = TRUE, echo = FALSE, fig.cap="\\label{fig:fig3}Histogram of dissolved organic carbon (mg/L)."}
#histogram to look at the distribution of doc
ggplot(carbon.data.skinny) +
  geom_histogram(aes(x = doc), 
                 binwidth = 0.5, fill = "grey", color = "black") +
  labs(x = "Dissolved Organic Carbon (mg/L)", y = "Count of DOC")

```

\autoref{fig:fig3} shows a histogram of the distribution of dissolved organic carbon (mg/L). From \autoref{fig:fig3} it is apparent that the distribution of dissolved organic carbon is not noramlly distributed. This information will factor into the statistical analyses conducted since the data is not normally distributed. 

```{r, eval = TRUE, echo = FALSE, fig.cap="\\label{fig:fig4}Frequency Polygon showing the count of DOC (mg/L) amounts in each Lake"}
#frequency polygon graph looking at DOC in each lake
ggplot(carbon.data.skinny) + 
  geom_freqpoly(aes(x = carbon.data.skinny$doc, color = Lake.Name), bins = 20) +
  labs(x = "Dissolved Organic Carbon (mg/L)", "Count of DOC", color = "Lake Name") +
  scale_color_brewer(palette = "Dark2")

```

\autoref{fig:fig4} shows the cumulative frequency of dissolved organice carbon in each lake. From \autoref{fig:fig4}, it is apparent that Paul lake has the highest cumulative frequency of dissolved organic carbon, followed by Peter lake. \autoref{fig:fig4} also shows that Peter and Paul lake mainly have dissolved organic carbon amounts between 0 mg/L and 10 mg/L, which is why there are two high spikes. The other lakes have dissolved orgnic carbon amounts that are more spread out, as shown in \autoref{fig:fig4}.  

```{r, echo=FALSE, eval=TRUE, fig.cap="\\label{fig:fig5} Dissolved Organic Carbon varying with depth. Depth increases from left to right, with Epilimnion being the surface of the lake."}
#graph looking at doc with depth
ggplot(carbon.data.skinny) +
  geom_boxplot(aes(x = depth, y = doc, fill = depth)) +
  labs(x = "Depth", y = "Dissolved Organic Carbon (mg/L)") +
  scale_fill_brewer("ece7f2", "a6bddb", "2b8cbe")

```

\autoref{fig:fig5} shows a box plot of each depth category of the lake against the doc measurements. From \autoref{fig:fig5} it looks like there is no relationship between depth and dissolved organic carbon (mg/L). \autoref{fig:fig5} indicates that the Epilimnion (water closest to the surface) and Hypolimnion (water from deepest part of the lake) have greater medians than the Metalimnion or the pooled mixed layer (PML). However, the PML has the greatest distribution of dissolved organic carbon, with one sample that might be considered an outlier as seen in \autoref{fig:fig5}.


```{r, echo=FALSE, eval=TRUE, fig.height = 12, fig.width = 10, fig.cap="\\label{fig:fig6}Distribution of dissolved organic carbon (mg/L) in each depth faceted by lake."}

#Violin plot with depth as the x-axis and doc as the y-axis
ggplot(carbon.data.skinny) + 
  geom_violin(aes(x = depth, y = doc, fill = depth), draw_quantiles = c(0.25, 0.5, 0.75)) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  labs(x = "Depth", y = "Dissolved Organic Carbon (mg/L)", fill = "Depth") +
  facet_wrap(vars(Lake.Name), ncol = 2) + 
  theme(panel.border = element_rect(color = "black", fill=NA, size=2), axis.text.x  =             element_text(size=12))
  
```

\autoref{fig:fig6} shows the distribution of dissolved organic carbon (mg/L) in each depth layer faceted by each lake. From \autoref{fig:fig6}, the PML of Peter lake has the greatest distribution of dissolved organic carbon (mg/L). Further, the distribution of dissolved organic carbon (mg/L) in each lake does not vary with depth. 


```{r, echo=FALSE, eval=TRUE, fig.height=6, fig.width=8, fig.cap="\\label{fig:fig7} Dissolved organic carbon (mg/L) over time in each lake."}
#plot looking at how doc varies over the years
ggplot(carbon.data.skinny, aes(x = Year, y = doc, color = Lake.Name)) +
  geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  scale_color_brewer(palette = "Set2") 
  
```

\autoref{fig:fig7} shows dissolved organic carbon over the years. \autoref{fig:fig1} was created to determine if there was a pattern of dissolved organic carbon in lakes over the years. From \autoref{fig:fig7} only Peter, Paul, and West long lake were continuously sampled very year. Other lakes, such as Hummingbird lake, were only sampled between the late 1990s to the early 2000s.

```{r, echo=FALSE, eval=TRUE, fig.width=8, fig.height=6, fig.cap="\\label{fig:fig8} Dissolved organic carbon (mg/L) by day of the year."}
#graph looking at how doc varies over time with Day Number on the x-axis
ggplot(carbon.data.skinny, aes(x = Day.Number, y = doc, color = Lake.Name)) +
  geom_point(alpha = 0.5) +
  labs(x = "Day of Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  scale_color_brewer(palette = "Set2")

```

\autoref{fig:fig8} shows the day number of the year plotted against dissolved organic carbon (mg/L) with different colors for each lake. From \autoref{fig:fig8} it is apparent that sampling was done mainly between days 140 and 250 of the year, which is during spring and summer months. Since samples were taken during the spring and summer months, it will not be possible to analyze whether DOC varies with depth seasonally.

\newpage

# Analysis

```{r}

#Shapiro-Wilkes test for normality
shapiro.test(carbon.data.skinny$doc)

#Bartlett test for equal variance
bartlett.test(carbon.data.skinny$doc, carbon.data.skinny$depth, 
              carbon.data.skinny$Lake.Name)

#Interaction effects with two-way ANOVA
carbon_aov <- aov(data = carbon.data.skinny, doc ~ depth*Lake.Name)
summary(carbon_aov)

#Interaction effects with two-way ANOVA with log of DOC
carbon_aov_log <- aov(data = carbon.data.skinny, log(doc) ~ depth*Lake.Name)
summary(carbon_aov_log) 

#post-hoc Tukey test using the glht function
K <- diag(length(coef(carbon_aov)))[-1,] 
rownames(K) <- names(coef(carbon_aov))[-1]

carbon.data.skinny$dl <- with(carbon.data.skinny, interaction(depth, Lake.Name))
cell <- aov(doc ~ dl - 1, data = carbon.data.skinny)
carbon.tukey <- summary(glht(cell, linfct = K))
carbon.tukey
```


```{r, echo=FALSE, eval=TRUE, fig.height=6, fig.width=5, fig.cap="\\label{fig:fig9}Plot of Post-Hoc Tukey test showing which pairwise differences are significant."}
#plot of post-hoc Tukey test
plot(carbon.tukey, cex.axis=0.45)

```

An interaction effects two-way anova test was run to determine if the interaction among lake name and depth is significant. The assumptions are that each variable had been taken independently of the other, that the data was taken from a normally distributed population, and that the variances in each group are equal. From \autoref{fig:fig1}, the data shown is not normally distributed, which thus violates the assumption of normality. Further, a Shapiro-Wilkes test was run to evaluate whether the dissolved organic carbon data is well approximated by a normal distribution. The Shapiro-Wilkes test showed that the data is not well approximated by a normal distribution (p < 2.2e-16). A Bartlett test was run on dissolved organic carbon, lake name, and depth to see if the variances were equal or different. The Bartlett test had a p-value less than 0.05, thus the null hypothesis is rejected and the alternate hypothesis that at least two of the variances are not equal.

From the results, the p-value for depth:Lake.Name is less than 0.05 (p-value = <2.e16), thus indicating that the interaction between these terms is siginifcant. Additionally the main effect, dissolved organic carbon, is significant. 

A non-parametric Post-Hoc Tukey test was run to determine the pairwise differences for the interaction since the interaction was significant. Pairs are in the same group if the p-value is greater than 0.05. From the Tukey test, the depths and lake names are all significantlly different. This is further evidenced in \autoref{fig:fig9}. 


```{r}
#wrangling dataset down to just the three lakes
Carbon.three.lakes <- 
  carbon.data.skinny %>%
  filter(depth == "Hypolimnion") %>%
  filter(Lake.Name == "Peter Lake" | Lake.Name == "Paul Lake" |
           Lake.Name == "West Long Lake")

#splitting data sets up into each lake data
carbon.peter2 <- filter(Carbon.three.lakes, Lake.Name == "Peter Lake")

carbon.paul2 <- filter(Carbon.three.lakes, Lake.Name == "Paul Lake")

carbon.west2 <- filter(Carbon.three.lakes, Lake.Name == "West Long Lake")
```

A Mann-Kendall test, which is a non-parametric trend test, was run to determine if there is a monotonic trend in dissolved organic carbon in Peter, Paul, and West Long lakes over time. This test was chosen because there is not a linear trend in DOC over time. These lakes were chosen based off of the \autoref{fig:fig7} which shows which lakes were monitored continuously over time. The hypolimnion was chosen because organic carbon can come from decomposed animals and plants in a lake which could settle at the bottom of the lake. Assumptions of a Mann-Kendall test include that methods for collecting data are unbiased, if there is no trend, the data are independently and identically distributed, and the measurements represent the true states of the observations when they were taken. In addition to a Mann-Kendall test, Pettitt's test were run. Pettitt test is a non-parametric test that can help determine if there is a change in tendancy and if so, where that change occurs (known as the changepoint).


```{r, eval=TRUE, echo=FALSE, fig.height = 6, fig.width = 8}
#plot of doc over time in the five selected lakes for just the hypolimnion
carbon.three.lakes.plot <- ggplot(Carbon.three.lakes, aes(x = Date, y = doc, color = Lake.Name)) + 
  geom_point(alpha=0.5, size = 1) +
  labs(x = "Date", y = "DOC (mg/L)", color = "Lake Name", title = "Peter, Paul, and West Long Lakes") +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "top")
 
```

```{r}
#Mann-Kendall test for dissolved organic carbon in each lake

#mk.test in Paul Lake
mk.test(carbon.paul2$doc) #high p-value so we accept the null that
#there is no trend in data
```

The Mann-Kendall test from Paul Lake shows that there is no trend in the data since the p-value is greater than 0.05 (p-value = 0.1589).

```{r, eval=TRUE, echo=FALSE, fig.height = 6, fig.width = 8}
#plot of Paul lake with changepoints in doc 
paul.lake <- ggplot(carbon.paul2, aes(x = Date, y = doc)) + 
  geom_point(alpha=0.5, size = 1, color = "blue") +
  labs(x = "Year", y = "DOC (mg/L)", color = "Lake Name", title = "Paul Lake") 
```


```{r}
#peter lake
mk.test(carbon.peter2$doc) #low p-value so reject the null and accept that there is a trend in data

#pettitt test to see where changepoint is
pettitt.test(carbon.peter2$doc) #changepoint is at 167

#mk.test before and after changepoint point
mk.test(carbon.peter2$doc[1:166]) #0.06 is greater than 0.05, 
#so accept that there is no trend in this part of the data
mk.test(carbon.peter2$doc[167:289]) #small p-value so could see a trend here

#pettitt test to see if there is a change point between 167 and 289
pettitt.test(carbon.peter2$doc[167:289]) #changepoint at 167+34 = 201

#mk.test between 167 and 200 and 201 and 289
mk.test(carbon.peter2$doc[167:200]) #p-value less than 0.05, 
#so might be another trend point in this area
mk.test(carbon.peter2$doc[201:289]) #p-value greater than 0.05, 
#so no trend point in this area

#pettitt test between 167 and 200
pettitt.test(carbon.peter2$doc[167:200]) #changepoint at 18+167 = 185

#mk.test between 167:184 and 185:200
mk.test(carbon.peter2$doc[167:184]) #p-value is equal to 1, 
#so accept the null that there are no more trend points
mk.test(carbon.peter2$doc[185:200]) #p-value is greater than 0.05,
#so accept the null that there are no more trends
```

The Mann-Kendall test for Peter Lake detected several trends in the data and the pettitt tets detected several change points. The changepoints are visible on \autoref{fig:fig10}. 

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 8}
#plot of Peter lake with changepoints in doc 
peter <- ggplot(carbon.peter2, aes(x = Date, y = doc)) + 
  geom_point(alpha=0.5, size = 1, color = "blue") +
  labs(x = "Year", y = "DOC (mg/L)", color = "Lake Name", title = "Peter Lake") +
  geom_vline(xintercept = as.Date("2007-08-14", origin = "1970-01-01"), color="#253494", lty = 2) + 
  geom_vline(xintercept = as.Date("2009-05-18", origin = "1970-01-01"), color="#253494", lty = 2) +  
  geom_vline(xintercept = as.Date("2010-05-31", origin = "1970-01-01"), color="#253494", lty = 2) 
```


```{r}
#Mann-Kendall tests and pettitt tests on West Long Lake to detect trends
mk.test(carbon.west2$doc)
pettitt.test(carbon.west2$doc) #changepoint at 60

mk.test(carbon.west2$doc[1:59]) #trend detected
mk.test(carbon.west2$doc[60:102]) #trend detected

#pettitt test
pettitt.test(carbon.west2$doc[1:59]) #changepoint at 16 + 1 = 17
pettitt.test(carbon.west2$doc[60:102]) #changepoint at 23+60 = 83
 
#mk.test on carbon west
mk.test(carbon.west2$doc[1:16]) #no trend
mk.test(carbon.west2$doc[17:59]) #trend detected

mk.test(carbon.west2$doc[60:82]) #no trend
mk.test(carbon.west2$doc[83:102])#trend detected 

#pettitt test on carbon west
pettitt.test(carbon.west2$doc[17:59]) #changepoint at 32 + 17 = 49
pettitt.test(carbon.west2$doc[83:102]) #changepoint at 7 + 83 = 90

#mk.test on West Long Lake
mk.test(carbon.west2$doc[17:48]) #no trend
mk.test(carbon.west2$doc[49:59]) #no trend

mk.test(carbon.west2$doc[83:89]) #trend detected
mk.test(carbon.west2$doc[90:102]) #no trend detected

#pettitt test on West Long lake
pettitt.test(carbon.west2$doc[83:89]) #changepoint at 2+83 = 85

#mk.test on West Long Lake
#mk.test(carbon.west2$doc[83:84]) #this cannot be run because 
#there must be at least 3 elements
mk.test(carbon.west2$doc[83:89])

```

The Mann-Kendall test for West Long Lake detected several trends in the data and the pettitt tets detected several change points. The changepoints are visible on \autoref{fig:fig10}. 

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 8, warning = FALSE, fig.cap="\\label{fig:fig8} DOC (mg/L) over time for West Long Lake."}
#plot of West Long lake with changepoints in doc 
west.long <- ggplot(carbon.west2, aes(x = Date, y = doc)) + 
  geom_point(alpha=0.5, size = 1, color = "blue") +
  labs(x = "Year", y = "DOC (mg/L)", color = "Lake Name", title = "West Long Lake") +
  geom_vline(xintercept = as.Date("1996-05-29", origin = "1970-01-01"), color="#253494", lty = 2) + 
  geom_vline(xintercept = as.Date("1998-06-17", origin = "1970-01-01"), color="#253494", lty = 2) +  
  geom_vline(xintercept = as.Date("1998-08-31", origin = "1970-01-01"), color="#253494", lty = 2) +
  geom_vline(xintercept = as.Date("2004-05-27", origin = "1970-01-01"), color="#253494", lty = 2) +
  geom_vline(xintercept = as.Date("2004-06-09", origin = "1970-01-01"), color="#253494", lty = 2) +
  geom_vline(xintercept = as.Date("2004-07-14", origin = "1970-01-01"), color="#253494", lty = 2) 
```

```{r, fig.height = 8, fig.width = 10, warning = FALSE, fig.cap="\\label{fig:fig10} DOC (mg/L) over time for Peter, Paul, and West Long Lakes. Changepoints in the data are represented by vertical lines."}
grid.arrange(carbon.three.lakes.plot, peter, west.long, paul.lake,
            nrow = 2)
```


\newpage

# Summary and Conclusions
<Summarize your major findings from your analyses. What conclusions do you draw from your findings? Make sure to apply this to a broader application for the research question you have answered.>

The significant findings from this analysis show that dissolved organic carbon does not vary by depth, as shown in \autoref{fig:fig5}. Instead, DOC was found to vary by lake, likely because each lake has a different depth and catchment area, which can greatly contribute to the amount of dissolved organic carbon found in the lake. Additionally, seasonal trends were not able to be determined between depth and DOC because DOC was measured primarily during the spring and summer months.

The two-way ANOVA with interactions test shows that the interaction between depth and lake name is significant. The post-hoc Tukey test indicates that the interaction between all of the lakes and depths are significant. A non-parametric Mann-Kendall test was conducted to see how DOC varies over time. From the Mann-Kendall test on dissolved organic carbon in the hypolimnion, Paul lake had no changepoints. This might have to do with the fact that Paul lake has the smallest range of dissolved organic carbon (mg/L) compared to Peter and West Long lakes. \autoref{fig:fig10} shows that Peter lake had three change points, where as West Long lake had six change points. West Long Lake had the most change points and this could be because it is 27m deep at its deepest, which is significantly deeper than Peter and Paul lakes, with 19.3m and 12.2m, respectively. Paul lake does not have any detected changepoints, which is likely because it was used as a control during an experiment on nutrients. Since nutrients were not added to Paul lake, the main source of DOC would be from the catchment area. However, nutrients were added to Peter lake which resulted in algal blooms. Algae can lead to an increase in DOC in lakes, which is likely why Peter lake has three changepoints. 