---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: Dissolved Organic Carbon in Wisconsin Lakes
subtitle: https://github.com/cwatson1013/Env_Data_Analysis_Final_Proj.git
author: Caroline Watson
abstract: "Dissolved organic carbon (DOC) is a significant part of the carbon cycle. DOC can come from within the lake, such as from decaying organisms or plant matter, or from the catchment area of the lake. The catchment area soils can influence the amount of DOC that enters each lake. The purpose of this study is to see if there is a relationship between depth and DOC and if so, if there is a seasonal relationship. This study looked at lakes that were part of the North Temperate Region located in Wisconsin, USA. DOC was found to not be normally distributed, so a two-way ANOVA with an interaction test was conducted. The results found that depth and lake name were significant. A non-parametric test was run on Peter, Paul, and West Long lakes to determine trends and changepoints in the data. Peter and West Long lake had trends and changepoints, but Paul lake did not. Overall, it was found that the depth and DOC do not have a relationship and there is no seasonal relationship between change in depth and DOC."
fontsize: 12pt
mainfont: Times New Roman
---

<Information in these brackets are used for annotating the RMarkdown file. They will not appear in the final version of the PDF document>

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

<Setup the global options for the R chunks in your document>

<Note: set up autoreferencing for figures and tables in your document>

```{r setup, include=FALSE, warning=FALSE, message=FALSE, echo=TRUE, eval=TRUE}

knitr::opts_chunk$set(echo = FALSE, eval = TRUE, fig.path = 'myproject/plot-',
                      cache = TRUE)

# Set your working directory
setwd("/Users/carolinewatson/Documents/Spring 2019/Environmental Data Analytics/Final_Project/Data/Code")

# Load your packages
suppressMessages(library(tidyverse))
library(ggplot2)
library(leaflet)
library(dplyr)
library(RColorBrewer)
library(viridis)
library(knitr)
library(kableExtra)
library(lubridate)
library(lme4)
library(nlme)
library(trend)
library(gridExtra)
library(multcomp)

# Set your ggplot theme
caroline_theme <- theme_classic(base_size = 16) + 
  theme(axis.text = element_text(color = "black"), 
        legend.position = "right")

theme_set(caroline_theme)
```


# Research Question and Rationale

<Paragraph detailing the rationale for your analysis. What is the significant application and/or interest in this topic? Connect to environmental topic(s)/challenge(s).>

The rationale for this analysis is that there typically is a relationship between dissolved organic carbon and depth. There is also typically a relationship between land area surrounding lakes that have high amounts of organic soils usually deposit large amounts of dissolved organic carbon into lakes. Dissolved organic carbon is an important part of the carbon cycle and supplies nutrients for some organisms. Most DOC is natural, but high amounts can indicate human influence, such as land surrounding the lake that is high in organic amount. 

<Paragraph detailing your research question(s) and goals. What do you want to find out? Include a sentence (or a few) on the dataset you are using to answer this question - just enough to give your reader an idea of where you are going with the analysis.>

This study aims to find out whether there is a relationship between dissolved organic carbon (DOC) and depth. If there is a relationship between these two variables, I want to see if this relationship varies seasonally. I am using a dataset that contains various parameter measurements for different lakes in the North Temperate Region in Wisconsin, USA. Parameters measured include temperature, depth, dissolved organic carbon, dissolved inorganic carbon, particulate organic matter and others.

\newpage

# Dataset Information

<Information on how the dataset for this analysis were collected, the data contained in the dataset, and any important pieces of information that are relevant to your analyses. This section should contain much of same information as the README file for the dataset but formatted in a way that is more narrative.>

<Add a table that summarizes your data structure. This table can be made in markdown text or inserted as a `kable` function in an R chunk. If the latter, do not include the code used to generate your table.>

```{r, message=FALSE, warning=FALSE, tbls="Summary of NTL-LTER Lake Carbon Data", results="asis"}

#reading in data file
carbon.data <- read.csv("../Raw/NTL-LTER_Lake_Carbon_Raw.csv")

#structure of data frame
carbon.data.summary <- summary(carbon.data)

#summary of data structure
kable(carbon.data.summary, caption = "Summary of Carbon Data from NTL-LTER Lakes in Wisconsin") %>% kable_styling(latex_options = c("hold_position", "striped", "scale_down"))

#install.packages("xtable")
library(xtable)
xtable(carbon.data.summary)

```

\newpage

# Exploratory Data Analysis and Wrangling

<Include R chunks for 5+ lines of summary code (display code and output), 3+ exploratory graphs (display graphs only), and any wrangling you do to your dataset(s).> 

```{r}
#class of sampledate column
class(carbon.data$sampledate)

#converting sampledate to a date in R
carbon.data$sampledate <- as.Date(carbon.data$sampledate, format = "%m/%d/%y")

#checking class of sampledate
class(carbon.data$sampledate)

#summary of the dataset
head(carbon.data)
summary(carbon.data)
colnames(carbon.data)
dim(carbon.data)

#renaming columns
colnames(carbon.data)[1:5] <- c("Lake.ID", "Lake.Name", "Year", "Day.Number", "Date")
```

```{r}
#filtering dataset by depth id and depth description and then selecting certain columns
carbon.data.skinny <- carbon.data %>%
  filter(depth_id == -2 | depth_id == -1 | depth_id == 7) %>%
  filter(depth == "Hypolimnion" | depth == "Epilimnion" | depth == "PML" |
           depth == "Metalimnion") %>%
  mutate(depth = factor(depth, levels = c("Epilimnion", "Metalimnion", "PML", "Hypolimnion"))) %>%
  filter(is.na(doc) == FALSE) %>%
  select(Lake.ID:depth_id, doc)

#converting dat column to date instead of factor
carbon.data$sampledate <- as.Date(carbon.data$Date, format = "%m/%d/%y")

#saving processed file to the processed folder
write.csv(carbon.data.skinny, 
          file = "../Processed/NTL-LTER_Lake_Carbon_Processed.csv")
```


```{r, echo=FALSE, eval=TRUE fig.height = 6, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig1} QQplot showing whether data for dissolved organic carbon is normally distributed."}
#qqplot of doc to see if there is a normal relationship
qqnorm(carbon.data.skinny$doc) 
qqline(carbon.data.skinny$doc)
```

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig2} Log QQplot of dissolved organic carbon data."}
#qqplot of log(doc) to see if this is better than the regular qqplot
qqnorm(log(carbon.data.skinny$doc))
qqline(log(carbon.data.skinny$doc))
```

\autoref{fig:fig1} is a qqplot that shows whether the dissolved organic carbon data has a normal distribution. From \autoref{fig:fig1}, it is clear that the dissolved organic carbon data do not follow a normal distribution. Becuase the data are not normally distributed, a log qqplot was run to see if log transforming the data would help the data look more normally distributed. \autoref{fig:fig2} shows the dissolved organic carbon data log transformed. From \autoref{fig:fig2} it is clear that dissolved organic carbon is not normally distributed, but log transforming the data makes the data more normally distributed than not log transforming the data. 

```{r, eval = TRUE, echo = FALSE, warning = FALSE, fig.cap="\\label{fig:fig3}Hisogram of dissolved organic carbon (mg/L)."}
#histogram to look at the distribution of doc
ggplot(carbon.data.skinny) +
  geom_histogram(aes(x = doc), 
                 binwidth = 0.5, fill = "grey", color = "black") +
  labs(x = "Dissolved Organic Carbon (mg/L)", y = "Count of DOC")

```

\autoref{fig:fig3} shows a histogram of the distribution of dissolved organic carbon (mg/L). From \autoref{fig:fig3} it is apparent that the distribution of dissolved organic carbon is not noramlly distributed. This information will factor into the statistical analyses conducted since the data is not normally distributed. 

```{r, eval = TRUE, echo = FALSE, warning = FALSE, fig.cap="\\label{fig:fig4}Dissolved Inorganic Carbon in each Lake"}
#frequency polygon graph looking at DOC in each lake
ggplot(carbon.data.skinny) + 
  geom_freqpoly(aes(x = carbon.data.skinny$doc, color = Lake.Name), bins = 20) +
  labs(x = "Dissolved Organic Carbon (mg/L)", "Count of DOC", color = "Lake Name") +
  scale_color_brewer(palette = "Dark2")

```

\autoref{fig:fig4} shows the cumulative frequency of dissolved organice carbon in each lake. From \autoref{fig:fig4}, it is apparent that Paul lake has the highest cumulative frequency of dissolved organic carbon, followed by Peter lake. \autoref{fig:fig4} also shows that Peter and Paul lake mainly have dissolved organic carbon amounts between 0 mg/L and 10 mg/L, which is why there are two high spikes. The other lakes have dissolved orgnic carbon amounts that are more spread out, as shown in \autoref{fig:fig4}.  

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig5} Dissolved Organic Carbon Varying with depth section of each lake. Depth increases from left to right in graph, with Epilimnion being the surface of the lake."}
#graph looking at doc with depth
ggplot(carbon.data.skinny) +
  geom_boxplot(aes(x = depth, y = doc, fill = depth)) +
  labs(x = "Depth", y = "Dissolved Organic Carbon (mg/L)") +
  scale_fill_brewer("ece7f2", "a6bddb", "2b8cbe")

```

\autoref{fig:fig5} shows a box plot of each depth category of the lake against the doc measurements. From \autoref{fig:fig5} it looks like there is no relationship between depth and dissolved organic carbon (mg/L). \autoref{fig:fig5} indicates that the Epilimnion (water closest to the surface) and Hypolimnion (water from deepest part of the lake) have greater medians than the Metalimnion or the pooled mixed layer (PML). However, the PML has the greatest distribution of dissolved organic carbon, with one sample that might be considered an outlier as seen in \autoref{fig:fig5}.


```{r, echo=FALSE, eval=TRUE, fig.height = 9, fig.width = 7, warning = FALSE, fig.cap="\\label{fig:fig6} Dissolved Organic Carbon concentrations in each lake. Depth of lake is represented by various colors."}

#Violin plot with depth as the x-axis and doc as the y-axis
ggplot(carbon.data.skinny) + 
  geom_violin(aes(x = depth, y = doc, fill = depth), draw_quantiles = c(0.25, 0.5, 0.75)) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  labs(x = "Depth", y = "Dissolved Organic Carbon (mg/L)", fill = "Depth") +
  facet_wrap(vars(Lake.Name), ncol = 2) + 
  theme(panel.border = element_rect(color = "black", fill=NA, size=2), axis.text.x  =             element_text(size=12))
  
```

\autoref{fig:fig6} shows the distribution of dissolved organic carbon (mg/L) in each depth layer faceted by each lake. From \autoref{fig:fig6}, the PML of Peter lake has the greatest distribution of dissolved organic carbon (mg/L). Further, the distribution of dissolved organic carbon (mg/L) in each lake does not vary with depth. 


```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 4.5, warning = FALSE, echo=FALSE, eval=TRUE, fig.cap="\\label{fig:fig7} Dissolved organic carbon concentration over time with year on the x-axis."}
#plot looking at how doc varies over the years
ggplot(carbon.data.skinny, aes(x = Year, y = doc, color = Lake.Name)) +
  geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  scale_color_brewer(palette = "Set2") 
  
```

\autoref{fig:fig7} shows dissolved organic carbon over the years. \autoref{fig:fig1} was created to determine if there was a pattern of dissolved organic carbon in lakes over the years. From \autoref{fig:fig7} only Peter, Paul, and West long lake were continuously sampled very year. Other lakes, such as Hummingbird lake, were only sampled between the late 1990s to the early 2000s.

```{r, echo=FALSE, eval=TRUE, fig.height = 6, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig8} Dissolved organic carbon concentration over time with day number of the year on the x-axis."}
#graph looking at how doc varies over time with Day Number on the x-axis
ggplot(carbon.data.skinny, aes(x = Day.Number, y = doc, color = Lake.Name)) +
  geom_point(alpha = 0.5) +
  labs(x = "Day of Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  scale_color_brewer(palette = "Set2")

```

\autoref{fig:fig8} shows the day number of the year plotted against dissolved organic carbon with different colors for each lake. From \autoref{fig:fig8} it is apparent that sampling was done mainly between days 140 and 250 of the year, which is during spring and summer months.  There are a cluseter of data points between 5mg/L and 15mg/L of dissolved organic carbon which indicates that there is likely not a linear relationship between dissovled organic carbon and time of year. Because of this, a time series analysis will be conducted as part of the statistical tests.

\newpage

# Analysis

```{r}

#Shapiro-Wilkes test for normality
shapiro.test(carbon.data.skinny$doc)

#Bartlett test for equal variance
bartlett.test(carbon.data.skinny$doc, carbon.data.skinny$depth, carbon.data.skinny$Lake.Name)

#Interaction effects with two-way ANOVA
carbon_aov <- aov(data = carbon.data.skinny, doc ~ depth*Lake.Name)
summary(carbon_aov) # use this anova and then try log transforming it; talk about assumptions of ANOVA and justify why moving forward with it; can use lme to get a list of all the interactions between depth and lakes 

carbon_aov_log <- aov(data = carbon.data.skinny, log(doc) ~ depth*Lake.Name)
summary(carbon_aov_log) #F value is lower with log transformed data than with data that is not log transformed

#post-hoc Tukey test
TukeyHSD(carbon_aov)

#post-hoc Tukey test using the glht function
comp <- glht(carbon_aov, linfct = mcp(depth = "Tukey"))
print(summary(comp))

comp.cld <- cld(comp)

carbon.data.skinny$dl <- with(carbon.data.skinny, interaction(depth, Lake.Name))
cell <- lm(doc ~ dl - 1, data = carbon.data.skinny)
summary(glht(cell, linfct = K))

#plot of post-hoc Tukey test
plot(comp.cld)

```

An interaction effects two-way anova test was run to determine if the interaction among lake name and depth is significant. The assumptions are that each variable had been taken independently of the other, that the data was taken from a normally distributed population, and that the variances in each group are equal. From \autoref{fig:fig1}, the data shown is not normally distributed, which thus violates the assumption of normality. Further, a Shapiro-Wilkes test was run to evaluate whether the dissolved organic carbon data is well approximated by a normal distribution. The Shapiro-Wilkes test showed that the data is not well approximated by a normal distribution (p < 2.2e-16). A Bartlett test was run on dissolved organic carbon, lake name, and depth to see if the variances were equal or different. The Bartlett test had a p-value less than 0.05, thus the null hypothesis is rejected and the alternate hypothesis that at least two of the variances are not equal.

From the results, the p-value for depth:Lake.Name is less than 0.05 (p-value = <2.e16), thus indicating that the interaction between these terms is siginifcant. Additionally the main effect, dissolved organic carbon, is significant. 

A non-parametric Post-Hoc Tukey test was run to determine the pairwise differences for the interaction since the interaction was significant. Pairs are in the same group if the p-value is greater than 0.05. From the Tukey test, a majority of the means of XXXXXXXXXXXX


```{r}
#wrangling dataset down to just the three lakes
Carbon.three.lakes <- 
  carbon.data.skinny %>%
  select(-Lake.ID, -depth_id) %>%
  filter(depth == "Hypolimnion") %>%
  filter(Lake.Name == "Peter Lake" | Lake.Name == "Paul Lake" | Lake.Name == "West Long Lake")

#splitting data sets up into each lake data
carbon.peter2 <- filter(Carbon.five.lakes2, Lake.Name == "Peter Lake")

carbon.paul2 <- filter(Carbon.five.lakes2, Lake.Name == "Paul Lake")

carbon.west2 <- filter(Carbon.five.lakes2, Lake.Name == "West Long Lake")
```

A Mann-Kendall test was run to determine if there is a monotonic trend in dissolved organic carbon in Peter, Paul, and West Long lakes over time. These lakes were chosen based off of the \autoref{fig:fig7} which shows which lakes were monitored continuously over time. The hypolimnion was chosen because organic carbon can come from decomposed animals and plants in a lake which could settle at the bottom of the lake, or the hypolimnion.


```{r}
#plot of doc over time in the five selected lakes for just the Metalimnion
carbon.five.lakes <- ggplot(Carbon.five.lakes2, aes(x = Date, y = doc, color = Lake.Name)) + 
  geom_point(alpha=0.5, size = 1) +
  labs(x = "Date", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  scale_color_brewer(palette = "Set2")
  
```

```{r}
#Mann-Kendall test for dissolved organic carbon in each lake

#mk.test in Paul Lake
mk.test(carbon.paul2$doc) #high p-value so we accept the null that there is no trend in data
```

The Mann-Kendall test from Paul Lake shows that there is no trend in the data since the p-value is greater than 0.05 (p-value = 0.1589).

```{r}
#plot of Paul lake with changepoints in doc 
paul.lake <- ggplot(carbon.paul2, aes(x = Date, y = doc)) + 
  geom_point(alpha=0.5, size = 1, color = "blue") +
  labs(x = "Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") 

print(paul.lake)
```


```{r}
#peter lake
mk.test(carbon.peter2$doc) #low p-value so reject the null and accept that there is a trend in data

#pettitt test to see where changepoint is
pettitt.test(carbon.peter2$doc) #changepoint is at 167

#mk.test before and after changepoint point
mk.test(carbon.peter2$doc[1:166]) #0.06 is greater than 0.05, so accept that there is no trend in this part of the data
mk.test(carbon.peter2$doc[167:289]) #small p-value so could see a trend here

#pettitt test to see if there is a change point between 167 and 289
pettitt.test(carbon.peter2$doc[167:289]) #changepoint at 167+34 = 201

#mk.test between 167 and 200 and 201 and 289
mk.test(carbon.peter2$doc[167:200]) #p-value less than 0.05, so might be another trend point in this area
mk.test(carbon.peter2$doc[201:289]) #p-value greater than 0.05, so no trend point in this area

#pettitt test between 167 and 200
pettitt.test(carbon.peter2$doc[167:200]) #changepoint at 18+167 = 185

#mk.test between 167:184 and 185:200
mk.test(carbon.peter2$doc[167:184]) #p-value is equal to 1, so accept the null that there are no more trend points
mk.test(carbon.peter2$doc[185:200]) #p-value is greater than 0.05, so accept the null that there are no more trends
```

The Mann-Kendall test for Peter Lake detected several trends in the data and the pettitt tets detected several change points. The changepoints are visible on \autoref{fig:fig_}. 

```{r, echo=FALSE, eval=TRUE}
#plot of Peter lake with changepoints in doc 
peter <- ggplot(carbon.peter2, aes(x = Date, y = doc)) + 
  geom_point(alpha=0.5, size = 1, color = "blue") +
  labs(x = "Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  geom_vline(xintercept = as.Date("2007-08-14", origin = "1970-01-01"), color="#253494", lty = 2) + 
  geom_vline(xintercept = as.Date("2009-05-18", origin = "1970-01-01"), color="#253494", lty = 2) +  
  geom_vline(xintercept = as.Date("2010-05-31", origin = "1970-01-01"), color="#253494", lty = 2) 

print(peter)
```


```{r, echo=TRUE, eval=TRUE}
#Mann-Kendall tests and pettitt tests on West Long Lake to detect trends
mk.test(carbon.west2$doc)
pettitt.test(carbon.west2$doc) #changepoint at 60

mk.test(carbon.west2$doc[1:59]) #trend detected
mk.test(carbon.west2$doc[60:102]) #trend detected

#pettitt test
pettitt.test(carbon.west2$doc[1:59]) #changepoint at 16 + 1 = 17
pettitt.test(carbon.west2$doc[60:102]) #changepoint at 23+60 = 83
 
#mk.test on carbon west
mk.test(carbon.west2$doc[1:16]) #no trend
mk.test(carbon.west2$doc[17:59]) #trend detected

mk.test(carbon.west2$doc[60:82]) #no trend
mk.test(carbon.west2$doc[83:102])#trend detected 

#pettitt test on carbon west
pettitt.test(carbon.west2$doc[17:59]) #changepoint at 32 + 17 = 49
pettitt.test(carbon.west2$doc[83:102]) #changepoint at 7 + 83 = 90

#mk.test on West Long Lake
mk.test(carbon.west2$doc[17:48]) #no trend
mk.test(carbon.west2$doc[49:59]) #no trend

mk.test(carbon.west2$doc[83:89]) #trend detected
mk.test(carbon.west2$doc[90:102]) #no trend detected

#pettitt test on West Long lake
pettitt.test(carbon.west2$doc[83:89]) #changepoint at 2+83 = 85

#mk.test on West Long Lake
mk.test(carbon.west2$doc[83:84])
mk.test(carbon.west2$doc[83:89])

```

The Mann-Kendall test for West Long Lake detected several trends in the data and the pettitt tets detected several change points. The changepoints are visible on \autoref{fig:fig_}. 

```{r, echo=FALSE, eval=TRUE}
#plot of West Long lake with changepoints in doc 
west.long <- ggplot(carbon.west2, aes(x = Date, y = doc)) + 
  geom_point(alpha=0.5, size = 1, color = "blue") +
  labs(x = "Year", y = "Dissolved Organic Carbon (mg/L)", color = "Lake Name") +
  geom_vline(xintercept = as.Date("1996-05-29", origin = "1970-01-01"), color="#253494", lty = 2) + 
  geom_vline(xintercept = as.Date("1998-06-17", origin = "1970-01-01"), color="#253494", lty = 2) +  
  geom_vline(xintercept = as.Date("1998-08-31", origin = "1970-01-01"), color="#253494", lty = 2) +
  geom_vline(xintercept = as.Date("2004-05-27", origin = "1970-01-01"), color="#253494", lty = 2) +
  geom_vline(xintercept = as.Date("2004-06-09", origin = "1970-01-01"), color="#253494", lty = 2) +
  geom_vline(xintercept = as.Date("2004-07-14", origin = "1970-01-01"), color="#253494", lty = 2) 

print(west.long)
```

```{r, fig.height = 4, fig.width = 6, warning = FALSE, fig.cap="\\label{fig:fig8} Dissolved organic carbon concentration over time with day number of the year on the x-axis."}
lay <- rbind(c(1,1,2,3),
             c(1,1,4,5),
             c(6,7,8,9))

grid.arrange(carbon.five.lakes, peter, west.long, paul.lake,
            nrow = 2)
             
#layout_matrix = lay)

```


\newpage

# Summary and Conclusions
<Summarize your major findings from your analyses. What conclusions do you draw from your findings? Make sure to apply this to a broader application for the research question you have answered.>

The major findings from this analysis show that dissolved organic carbon does not vary by depth, but rather by lake. This is likely due to the fact that each lake has a different depth and catchment area, which can greatly contribute to the amount of dissolved organic carbon found in the lake. The two-way ANOVA with interactions test shows that the interaction between depth and lake name is significant. The post-hoc Tukey test indicates that the interaction between a majority of lakes and depths are significant. From the Mann-Kendall test on dissolved organic carbon in the hypoliminion, Paul lake had no changepoints. This might have to do with the fact that Paul lake has the smallest range of dissolved organic carbon (mg/L) compared to Peter and West Long lakes. \autoref{fig:fig_} shows that Peter lake had three change points. \autoref{fig:fig_} shows that West Long lake had six change points. 

Overall, the analyses determined that a seasonal trend between depth and dissolved organic carbon was not possible based off the data available since most of the samples were conducted during the spring and summer months. Because \autoref{fig:fig5} shows the distribution of dissolved organic carbon (mg/L) in each depth layer and there is no clear trend. Although the original research question was to look at the relationship between depth and dissolved organic carbon, the exploratory data analysis showed that there was no relationship. Because there was no relationship between dissolved organic carbon and depth, the focus of this study had to change. Each lake has a different depth, so instead of running a generalized linear model, a two-way interaction effects anova was run with depth and lake name as the interaction terms. 

Figure ___ (daynumber and DOC concentration figure) shows that sampling was only done in the summer, which indicates that it will not be possible to detect a seasonal trend between depth and DOC. Further, Figure _ (qqplot figures) show that the dissolved organic carbon data is not normally distributed, thus leading to the conclusion that a linear model will not be feasible with this data. When the dissolved oranic carbon is log transformed as Figure (log transformed qqplot) shows, the data looks more normal than it did when not log transformed. Also, figure (violin plot figure) shows that DOC varies accross the different lakes, which leads to the anova test where Lake is included as an interaction with depth. Also, it is apparent from figure (boxplot of depth) that DOC does not increase or decrease with depth; its most likely variable, and thus likely variable amongst lakes.  

Peter Lake has a high concentration likely due to the fact that there was an experiment with Peter and Paul lakes, where nutrients were put into the lakes. 
